Model Checking for Open Concurrent Systems

### 2018-09-12 2018-ModelCheckingPNets/checking.tex

\let\springervec\vec
\let\vec\origvec
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{latexsym}
\setcounter{tocdepth}{3}

@@

\begin{figure}
	\centering
	\begin{tikzpicture}[->, shorten >=2pt, >=stealth, node distance=3cm,
	noname/.style={ellipse, minimum width=5em, minimum height=3em,draw}]
	\node[noname] (1)                                             {$s_{0}$};
	\node[noname] (2) [right=of 1]                                {$s_{1}$};

	\path 	(1) edge [bend left]                  node[above] {$[true],\langle start?M\rangle,\{t:=M\}$} (2)
	(2) edge [bend left]                  node[below] {$[t=0], \langle timeout\rangle$} (1)
	(2) edge 			               node[above] {$\langle resume\rangle$} (1);
	\path[->] (2) edge [in=-10,out=20,loop] coordinate[pos=0.52] (midp) node[right] {$[t>0], \langle tick\rangle, \{t:=t-1\}$} (2);

	\end{tikzpicture}
	\caption{The pLTS of the Timer component}
\end{figure}

\subsubsection{Semantics of pLTS}

@@

\begin{table}
	%	\multirow{2}{*}{Algorithme}
	\setlength\tabcolsep{18pt}% default value is 6pt 
	\centering 
	\caption{Syntax of the pM$\mu$} 
	\label{mcl}
	\begin{tabular}{>{\bfseries}c| c } 
		\hline 		
	Action formula 	&     $\alpha ::= \{\{c(arg_{1}, ..., arg_{n}), \tau\}\mid arg_{i}=?x\ |\ !e\  \}$                  \\
	& $|\ \neg\alpha \ |\ \alpha_{1}\vee\alpha_{2}\ |\ \alpha_{1}\wedge\alpha_{2}$   \\ 	\hline

@@

 where, $c$ is an action name expressing the communication port. $arg_{i}$ is a list of arguments, and each $arg_{i}$ is either a send expression $!e$, or a receive variable $?x$. An action $\alpha$ is a parameterised action $c(arg_{i})$ or a silent action $\tau$. $p$ is an atomic proposition and $X$ is a set of variables. $\neg$, $\vee$ and $\wedge$ are respectively negation, disjunction and conjunction operator on action formula and state formula. $[\alpha]\varphi$ expresses "whenever an action $\alpha$ happens, we will do $\varphi$", and $\langle\alpha\rangle$ expresses "an action $\alpha$ takes effective, we will do $\varphi$". $x_{i}\in X$ is a free variable bound in state formula $\varphi$. $\mu$ and $\nu$ are respectively the minimum fixed point and the maximum fixed point.\TODO{Tengfei: Must more explaination}\NOTE{Done}

We employ the case study from timer component\cite{mavridou2016architectrue}. A liveness property defines that good thing finally happens. In this case study, intuitively the timer finally returns to the initial state after some execution steps. The property can be expressed by a pM$\mu$ formula:

@@

\ERIC{this description is too operational (like a state machine), I would like a more logical description of the formula. One way is to use the vocabulary of ``Specification patterns'', like ``it is possible'', ``it is mandatory'', ``do something until some condition'', etc}
At first, we can do $\langle start?M\rangle$ and then it is mandatory to enters a loop with an assignment $t := M$. In the loop, it is possible for us to do a $timeout$ until  $t$ is equal to 0, and then it is done and the result is true. Or we can do a $tick$ and then we continue the loop with decrementing the value $t$ by one.

@@

\begin{figure}
    \centering\includegraphics[width=5.5in]{figures/architectrue.jpeg} 
    \caption{Architecture of symbolic model checking}
    \label{architectrue}
\end{figure}

@@

\subsection{The difference between pLTS and open automata}
 pLTS is a labelling transition system with parameterised action. Open pNet is a composition of some pLTSs with synchronisation vector and some holes. A BIP architecture of an open pNet includes coordinator components and some operand components. An open automaton is formed by the unfolding of an open pNet.

\subsection{}

\subsection{}

>>>>

\let\springervec\vec
\let\vec\origvec
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{latexsym}
\setcounter{tocdepth}{3}
\usepackage{mathpartir}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\columnwidth]{figures/TimerpLTS.eps}
  \caption{The pLTS of the Timer component}
  \label{schema:Timer:pLTS}
\end{figure}

@@

\begin{table}
	%	\multirow{2}{*}{Algorithme}
	\setlength\tabcolsep{18pt}% default value is 6pt 
	\centering 
	\caption{Syntax of the pM$\mu$} 
	\label{mcl}
	\begin{tabular}{>{\bfseries}c| c } 
		\hline 		
	Action formula 	&     $\alpha ::= \{c(arg_{1}, ..., arg_{n})\mid arg_{i}=?x\mid\ !e\mid \tau\}$                  \\
	& $|\ \neg\alpha \ |\ \alpha_{1}\vee\alpha_{2}\ |\ \alpha_{1}\wedge\alpha_{2}$   \\ 	\hline

@@

 where, $c$ is an action name expressing the communication port. $arg_{i}$ is a list of arguments, and each $arg_{i}$ is either a send expression $!e$, a receive variable $?x$ or a silent action $\tau$. An action $\alpha$ is a parameterised action $c(arg_{i})$. $p$ is an atomic proposition and $X$ is a set of variables. $\neg$, $\vee$ and $\wedge$ are respectively negation, disjunction and conjunction operator on action formula and state formula. $[\alpha]\varphi$ expresses "whenever an action $\alpha$ happens, we will do $\varphi$", and $\langle\alpha\rangle$ expresses "an action $\alpha$ takes effective, we will do $\varphi$". $x_{i}\in X$ is a free variable bound in state formula $\varphi$. $\mu$ and $\nu$ are respectively the minimum fixed point and the maximum fixed point.\TODO{Tengfei: Must more explaination}\NOTE{Done}

We employ the case study from timer component~\cite{mavridou2016architectrue}. A liveness property defines that good thing finally happens. In this case study, intuitively the timer finally returns to the initial state after some execution steps. The property can be expressed by a pM$\mu$ formula:

@@

\ERIC{this description is too operational (like a state machine), I would like a more logical description of the formula. One way is to use the vocabulary of ``Specification patterns'', like ``it is possible'', ``it is mandatory'', ``do something until some condition'', etc}

At first, the timer starts with receiving a value $M$, and then it is mandatory to enters a loop with an assignment $t := M$.  In the loop $\mu$, it is possible for the timer to return true until the time is out $\langle timeout\rangle$, and it is done. Or it will continue the loop with decreasing the value $t$ by one when the timer ticks again $\langle tick\rangle$.

@@

\begin{figure}
    \centering\includegraphics[width=5.5in]{figures/architecture.pdf}
    \caption{Architecture of symbolic model checking}
    \label{architectrue}
\end{figure}

@@

\subsection{Open Automata}

The semantics of open pNets will be defined  as an open automaton. An open
automaton is an automaton where each transition composes transitions of several LTSs with action of some holes, the transition occurs if some predicates hold, and can involve a set of state modifications.

\begin{definition}[Open transitions]
	\label{def:OpenTransitions}
	An \emph{open transition} over a set  $(S_i,s_{0 i}, \rightarrow_i)^{i\in
	I}$ of LTSs, a
	set $J$ of holes with sorts $Sort_j^{j\in J}$, and a set of states $\mathcal{S}$ is a structure of the form:	
	\begin{mathpar}
	\inferrule*[myfraction=\reddottedrule]
	{\{s_i~{\xrightarrow{a_i}}_i ~s_i^{\prime}\}^{i\in I},
		\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred, \Post}
	{s \xrightarrow {v}s'}
	\end{mathpar}
	Where $s, s'\in\mathcal{S}$ and for all
        $i\in I$, $s_i{\xrightarrow{a_i}}_i s_i^{\prime}$ is a transition of the
	LTS $(S_i,s_{0 i}, \rightarrow_i)$, and $\xrightarrow{b_j}_j$
        is a transition of the hole $j$, for any action $b_j$ in the
        sort $Sort_j$. \Pred\ is a predicate
	over the different variables of the
	terms, labels, and states $s_i$, $b_j$, $s$, $v$. \Post\ is a set of equations that 
	hold \emph{after the open transition}, they are represented as a substitution of the 
	form $\{x_k\gets e_k\}^{k\in K}$ 
	where $x_k$ are variables of $s'$, $s'_i$, and $e_k$ are expressions over the other 
	variables of the open transition.
\end{definition}

\begin{definition}[Open automaton]
	\label{def:open-automaton}
	An \emph{open automaton} is a structure\\ $A =
	<LTS_i^{i\in I},J,\mathcal{S},s_0,\mathcal{T}>$ where:
	\begin{itemize}
		\item[$\bullet$]  $I$ and $J$ are  sets of indices,
		\item[$\bullet$]  $LTS_i^{i\in I}$ is a family of LTSs,
		\item[$\bullet$]   $\mathcal{S}$ is a set of states and $s_0$ an initial state
		among $\mathcal{S}$,
		\item[$\bullet$] $\mathcal{T}$ is a set of open transitions and for each
		$t\in \mathcal{T}$ there exist $I'$, $J'$ with $I'\subseteq I$, $J'
		\subseteq J$, such that $t$ is an open transition over $LTS_i^{i\in I'}$, $J'$,
		and  $\mathcal{S}$.

	\end{itemize}
\end{definition}

\begin{definition}[States of open pNets]\label{def-states}
  A state of an open pNet is a tuple (not necessarily finite) of the
  states of its leaves (in which we denote tuples
  in structured states as $\triangleleft\ldots\triangleright$ for better readability).

  For any pNet p, let $\overline{Leaves} = \mylangle S_i,{s_i}_0, \to_i\myrangle^{i \in L}$ be the set of pLTS at its leaves,
  then $States(p) = \{\triangleleft s_i^{i\in L}
  \triangleright| \forall i\in L. s_i \in S_i\}$.
A pLTS being its own single leave:
  $States(\mylangle S,s_0, \to\myrangle) = \{\triangleleft s \triangleright| s \in S\}$.

The initial state is defined as:
$InitState(p) = \triangleleft {{s_i}_0}^{i\in L}  \triangleright$.
\end{definition}

\paragraph{Predicates:}
Let $\mylangle\overline{\pNet},\overline{S},\symb{SV}_k^{k\in K} \myrangle$
be a pNet. Consider a synchronization vector $SV_k$, for $k\in K$. We define a
predicate $\Pred$ relating the actions of the involved sub-pNets and the resulting actions. This predicate verifies:
\[\Pred(SV_k, a_i^{i\in I}, b_j^{j\in J}, v)\Leftrightarrow
\begin{array}{l}
\exists {(a'_i)}^{i\in I},
{(b'_j)}^{j\in J},v'.\, SV_k={(a'_i)}^{i\in I}, {(b'_j)}^{j\in J}\rightarrow v'
\\~~\land
\forall i\in I.\, a_i=a'_i\land \forall j \in J.\, b_j=b'_j \land v=v'
\end{array}\]
In any other case (if the action families do not match or if there is no valuation of variables such that the above formula can be ensured) the predicate is undefined.

This definition is not constructive but it is easy to build the predicate constructively by brute-force unification of the sub-pNets actions with the corresponding vector actions, possibly followed by a simplification step.

 \subsection{The difference between pLTS and open automaton}
 pLTS is a labelling transition system with parameterised action. An open pNet is a composition of some pLTSs and some holes, and a set of synchronization vector composites the action of an open pNet. A BIP architecture of an open pNet includes coordinator components and some operand components. An open automaton is formed by the unfolding of an open pNet through the rules in~\cite{xudong2017smt}.

 In order to check the model of open automata, the algorithm in the appendix have to be modified. There are two different things in the structure of open automaton and pLTS.
 \subsubsection{Hole}
 A hole in pNet is defined without states, but it involves in a transition, which is defined as $\{\xrightarrow{b_j}_j\}^{j\in J}$. The action $b_{j}$ is the action of a hole with an index $j$. The difference of holes from pLTS is whether a component involves in a set of states. However, each action in the hole belongs to a set of the parameterised action.

 \subsubsection{Predicate}
 A transition in a pLTS is consisted of a tuple with labelling $guard$, $parameterised\; action$, $assignment$ between two states. Formally, a transition is defined as $s\xrightarrow{\langle\alpha, e_{b}, (x_{j}:=e_{j})^{j\in J}\rangle}s'$. While an open transition in an open automaton is defined with labelling $Pred$ and $Post$ between a composition of states in pLTSs and holes. In the open transition, $Pred$ is a predicate over the different variables of the terms, labels, and states. Specifically, a predicate consists of a set of parameterised actions in pLTSs and actions in holes. In a pNet, some ports from each component have an immediate relationship, and these ports are connected each other. Also, one of these ports is treated as a action to trigger a transition in the pLTS or in a hole. A predicate is a synchronization vector from the action in each pLTS and each hole.

\subsection{opM$\mu$}
In order to check open automata, a new property language is present. The property language is called open parameterised modal $\mu$-calculus (opM$\mu$). We give the formal definition in Tab.~\ref{opmu}:

\begin{table}
	%	\multirow{2}{*}{Algorithme}
	\setlength\tabcolsep{18pt}% default value is 6pt 
	\centering 
	\caption{Syntax of the opM$\mu$} \label{opmu}
	\begin{tabular}{>{\bfseries}c| c } 
		\hline 		
	Action formula 	&     $\alpha ::= \{c(arg_{1}, ..., arg_{n})\mid arg_{i}=?x\mid\ !e\mid \tau\}$                  \\
	& $|\ \neg\alpha \ |\ \alpha_{1}\vee\alpha_{2}\ |\ \alpha_{1}\wedge\alpha_{2}$   \\ 	\hline

	Predicate 	&     $Pred ::= \{c_i(arg_{1}, ..., arg_{n})\mid arg_{j}=?x\mid\ !e\mid \tau\}$                  \\
	& $|\ \neg\alpha \ |\ \alpha_{1}\vee\alpha_{2}\ |\ \alpha_{1}\wedge\alpha_{2}$   \\ 	\hline

\subsection{The modified algorithm}
